An Ant Cuticle Texture Classification Algorithm for Ecological Anaylsis

### 2021-09-14 paper/paper.tex
jp: expand sculpturing sections

\section{Related Work}

@@

\subsection{Sculpture Identification Protocol}

@@

\begin{table}[h]
  \centering
  \caption{Dataset Subclass Distribution}
  \label{table:classdist_all_full}
  \include{tables/all_samples_full.tex}
\end{table}

\begin{table}[h]
  \centering
  \caption{Dataset Class Distribution}
  \label{table:classdist_rs_full}
  \include{tables/rs_samples_full.tex}
\end{table}

Images were collected and classified by undergraduate students according to the
sculpturing identification protocol. Overall, there are a total of 2443 images
in the dataset. The class distribution with all considered subclasses is shown
in Table \ref{table:classdist_all_full}, and the class distribution with overall
distribution is shown in Table \ref{table:classdist_rs_full}. The relatively
small number of images in the dataset paired with the uneven class distribution
makes it difficult to perform a meaningful classification. However, the goal of
this project is to develop a classification algorithm that can be used to help
classify over 200,000 ant images. Therefore, it is desirable to use a few
representative images from each class to train the algorithm. Additionally, data
augmentation and other methods are performed to increase the number of training
images to create a more robust classification algorithm.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{assets/images/segmented_1.png}
  \caption{Hand segmented ant head image to separate head cuticle from other
    sections of the image.
    \href{https://www.antweb.org/bigPicture.do?name=casent0217419&shot=h&number=1}{CASENT0217419}
    by \href{https://www.antweb.org/artist.do?id=92}{Estella Ortega}, from
    \href{https://www.antweb.org}{AntWeb}, is licensed under
    \href{https://creativecommons.org/licenses/by/4.0/}{CC BY 4.0}}
  \label{fig:segmented_1}
\end{figure}

@@

\begin{table}[h]
  \centering
  \caption{Subimage Dataset Class Distribution}
  \label{table:classdist_rs_sub}
  \include{tables/rs_samples_sub.tex}
\end{table}

@@

@@

\begin{table}[h]
  \centering
  \caption{Binary Classification of Full Size Images}
  \label{table:exp1_full}
  \include{tables/experiment_1.tex}
\end{table}

>>>>

\section{Related Work}
Taxonomists have developed extensive terminology describing ant cuticle
sculpturing as it is often a useful diagnostic trait to distinguish between
closely related ant species \cite{blaimer_taxonomy_2019,fisher_ants_2007}. The
definitive text on ant cuticle terminology — \textit{The Glossary of Surface
  Sculpturing} — contains over 100 terms to describe the cuticle sculpturing
patterns of ants \cite{harris_glossary_1979}. There is often substantial overlap
among terms, and closely related cuticle patterns are likely to be functionally
similar. For example, the definition for “imbricate” is, “partly overlapping and
appearing like shingles on a roof or scales on a fish,” which is difficult to
distinguish from the definition of tesselate, “made up of squares like a chess
board, either in sculpturing or in color.”

Cuticle sculpturing in ants has been explored thoroughly from a taxonomic
perspective; however, the function of nano and microstructures on insect
exoskeletons is a developing topic in entomology. Watson et al. reviews the
literature of cuticle nano and microstructure function and proposes 21 possible
functions associated with these structures \cite{watson_diversity_2017}. Many of
these functions are related to structures not found in ants such as scales and
nanostructures. The review does include functions that may relate to ants such
as friction control, enhanced surface area, and increased hardness.  Watson et
al. also describe seven types of cuticle structures ranging from hairs and
scales to nano and micro structures \cite{watson_diversity_2017}. The cuticle
sculpturing of ants seems to fall within one type – complex microstructures.

The five broad functional groupings developed that describes the complex
microstructures found on ants were derived from reviewing the variation of
cuticle sculpturing across ants. These functional groupings were also reflected
in Harris as many terms could be grouped together based on similar definitions
and comparing the SEM photographs provided in the publication
\cite{harris_glossary_1979}.

@@

\subsection{Sculpture Identification Protocol}
A team of three assistants were in charge of the manual identification process.
The team was trained to identify cuticle sculpturing through a process which
consisted of one 45-minute introductory lesson explaining the project and
texture categories and then given a training set of photos to identify from the
genus \textit{Polyrhachis}, whose members display among the highest diversity of
cuticle patterns all categories. The training set identifications were reviewed
together as a group. Once training was complete, assistants were assigned the
same genera of ants to identify independently each week. A weekly meeting was
held to discuss identifications and assign new ones. These identifications were
collected in a spreadsheet and the identifications were assigned to individual
ant species on a majority basis.

@@

\begin{table}[ht]
  \centering
  \caption{Dataset Subclass Distribution}
  \label{table:classdist_all_full}
  \include{tables/all_samples_full.tex}
\end{table}

\begin{table}[ht]
  \centering
  \caption{Dataset Class Distribution}
  \label{table:classdist_rs_full}
  \include{tables/rs_samples_full.tex}
\end{table}

Images were collected and classified by according to the sculpturing
identification protocol. Overall, there are a total of 2443 images in the
dataset. The class distribution with all considered subclasses is shown in Table
\ref{table:classdist_all_full}, and the class distribution with overall
distribution is shown in Table \ref{table:classdist_rs_full}. The relatively
small number of images in the dataset paired with the uneven class distribution
makes it difficult to perform a meaningful classification. However, the goal of
this project is to develop a classification algorithm that can be used to help
classify over 200,000 ant images. Therefore, it is desirable to use a few
representative images from each class to train the algorithm. Additionally, data
augmentation and other methods are performed to increase the number of training
images to create a more robust classification algorithm.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{assets/images/segmented_1.png}
  \caption{Hand segmented ant head image to separate head cuticle from other
    sections of the image.
    \href{https://www.antweb.org/bigPicture.do?name=casent0217419&shot=h&number=1}{CASENT0217419}
    by \href{https://www.antweb.org/artist.do?id=92}{Estella Ortega}, from
    \href{https://www.antweb.org}{AntWeb}, is licensed under
    \href{https://creativecommons.org/licenses/by/4.0/}{CC BY 4.0}}
  \label{fig:segmented_1}
\end{figure}

@@

\begin{table}[ht]
  \centering
  \caption{Subimage Dataset Class Distribution}
  \label{table:classdist_rs_sub}
  \include{tables/rs_samples_sub.tex}
\end{table}

@@

\subsection{Sculpturing Identification}
In our initial trial, we examined the entire \textit{Polyrhachis} genus, which
was 581 eligible specimens out of the 775 species. 71\% of the classifications
were unanimous. We then expanded our categorizations to 2846 images of
individual species from 13 genera. These genera were from the four major ant
subfamilies Ponerinae, Myrmicinae, Formicinae, and Dolichoderinae along with
Dorylinus and Ectatomminae. 87\% (2486) of classifications were unanimous. This
percentage represents the accuracy for the entirety of identifications starting
November 2020 and ending in March 2021. Assistants were also asked to complete
an ant cuticle assessment near the end of the identification process. The
assessment consisted of images from forty species of ants that spanned all
categories. The purpose of the assessment was to measure the consistency of
identifications while removing any biases from weekly meeting discussions and
cross referencing of other identifications. Unanimous classification was reached
90\% of the time during the individual assessment.

Classifications became more consistent overtime. One likely reason for this
increase was that students gained additional experience in cuticle
classification over the course of these trials. However, definitions of cuticle
texture categories were refined overtime. At the inception of the project,
assistants based classifications off of reference photos and tentative
definitions that were sometimes unclear. For example, striate was initially
defined as “long linear ridges, corresponding in-between ridges vary in length.”
Also, some tentative categories had low agreement amongst the team. For example,
striate initially was a broad category containing two subclasses (Severe and
Mild) and encompassed longitudinal and latitudinal ridges. Based on several
rounds of feedback from students and identification of these problem categories
where there was low agreement, the definitions developed into the current
classification system.

\subsection{Binary Classification}

@@

\begin{table}[ht]
  \centering
  \caption{Binary Classification of Full Size Images}
  \label{table:exp1_full}
  \include{tables/experiment_1.tex}
\end{table}
